---
title: "Exploratory Analysis"
output: html_document
---

## Project title

Give your project a concise, interesting title that summarizes the entirety of your project

### Authors

Names of your team members

### Date

Winter 2022

### Abstract

No more than three sentences that summarize your project. Focus on the very most important aspects. For example: (1) "Our main question is .... This question is important because .... To address the question, we will ...." (2) "We are concerned with ..., because .... To address this concern, we plan to ...." (3) "Consider that .... This is important because .... Accordingly, we plan to ...."

### Keywords

3-5 keywords that summarize your project.
(e.g., "Keywords: human physiology; bicycle exercise; Gen Z; times-series data")

### Introduction  

Briefly introduce your project.  Include 3-5 research questions. What motivates the questions? Why are they important? (at least 200 words)

### Related Work  

Describe your topic and related work in this space. You must include 3 citations to related work (URLs to similar work, high quality articles from the popular press, research papers, etc. ) Please use a standard citation style of your choice. (at least 200 words)

### The Dataset

Where did you find the data? Please include a link to the data source  
Who collected the data?  
How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)

### Implications

Assuming you answer your research questions, briefly describe the expected or possible implications for technologists, designers, and policymakers. (at least 150 words)

### Limitations & Challenges
What challenges or limitations might you need to address with your project idea more broadly? Briefly discuss. (at least 150 words)

### Summary Information

Write a summary paragraph of findings that includes the 5 values calculated from your summary information R script

### Table

Include a table of aggregate information

Describe why you included the table and what information it reveals

### Chart 1

Total fatalities from year 1908 to 2009

Describe why you chose this chart and what information it reveals

I chose to present this chart because I wanted to look at the data of total fatalities
from all the years the dataset provides. See which year had the highest fatality rate.

I found out that the highest fatalities during this timeframe was during 1972.
Total fatality count of 2937.

``````{r, echo = FALSE, code = readLines("chart1_JJ.R")}
```

### Chart 2

Top 20 Aircraft Makes with the most Deaths & Top 20 Aircraft Types with the most Deaths

Describe why you chose this chart and what information it reveals

I chose to put together these charts because I thought that it would be interesting to see what aircraft Makes resulted in the most deaths, as well as which specific make resulted in the most deaths.

I was able to find that Boeing has the most deaths of all Aircraft makes. Though the Aircraft type resulting in the most deaths the Douglas DC-3, is not produced by Boeing... which I found particularly interesting.

``````{r, echo = FALSE, code = readLines("chart2_NoahK.R")}
```

### Chart 3

Include a chart

Describe why you chose this chart and what information it reveals
